Graph Generators: State of the Art and Open Challenges

### 2019-06-14 generators_SocialNetworks.tex
updated social network generators strenghts and weaknesess

Another important aspect
of social networks is the community effect. A detailed study of structure of
communities in 70 real-world networks is provided, e.g., in
~\cite{Leskovec:2008:SPC:1367497.1367591}.
~\cite{Prat-Perez:2014:CSS:2621934.2621942} analyzed the structure of
communities (clustering coefficient, triangle participation ratio, bridges,
diameter, conductance and size) in both real-world graphs and outputs of existing graph
generators LFR~\cite{PhysRevE.78.046110} and the
LDBC-SNB~\cite{Erling:2015:LSN:2723372.2742786}. They found out that discovered communities  in different graphs have common distributions and that communities in a single graph have diverse nature and are difficult to fit with a single model.

@@

\paragraph{Realistic Social Network}
~\cite{Barrett:2009:GAL:1995456.1995598} focused on the construction of
realistic social networks. For this purpose the authors combine both public and private data sets
with large-scale agent-based techniques. The process works as follows: In the first step
it generates a synthetic data by combining public and commercial databases. In the second step, it determines a set of activity templates. A 24-hour activity sequence including
geolocations is assigned to each synthetic individual. To demonstrate the approach, the authors create a synthetic US population consisting of people and households together with respective geolocations. For this purpose the authors combine simulation and data fusion techniques utilizing various real-world data sources such as U.S. Census data, responses to an activity survey or a time-use survey. %Demographic information for each person and location, a minute-by-minute schedule of each person's activities, and the locations where these activities take place is generated by a combination of simulation and data fusion techniques.
The result is captured by a dynamic social contact network. Similar methods for agent-based strategies have been reported in~\cite{Bernstein:2013:SAS:2499604.2499609}.

\paragraph{Linkage vs. Activity Graphs} \cite{Yao2011} distinguished between two
types of social network graphs -- the \emph{linkage graph}, where nodes represent people and edges correspond to their friendships, and the \emph{activity graph}, where nodes also represent people but correspond to their interactions. On the basis of the analysis of
Flickr\footnote{\url{https://www.flickr.com/}} social links and
Epinions\footnote{\url{http://www.epinions.com/}} network of user interactions, the authors discover that they both exhibit high clustering coefficient (community structure), small diameter, and power-law degree distribution. Considering the dynamic properties they both have relatively stable clustering coefficient over time and follow the densification law. On the other hand, diameter shrinking is not observed in Epinions activity graph and there is a difference in degree
correlation (i.e., frequency of mutual connections of similar nodes) -- linkage graphs have positive degree correlation, activity graphs have neutral degree correlation. With regards to the findings, the proposed generator focusses on linkage graphs with positive degree correlation. For this purpose it extends the forest fire spreading process algorithm~\cite{Leskovec:2005:GOT:1081870.1081893} with link symmetry. It has two parameters: the \emph{burning probability} $P_b$ and the \emph{symmetry probability} $P_s$. $P_b$ ensures a BFS-based forward burning process in which fire burns strongly with $P_b$ approaching 1.  $P_s$ ensures backward linking from old nodes to new nodes and ``adds fuel to the fire as it brings more links''. %It gives chances for big nodes to connect back to big nodes.

@@

\paragraph{Cloning of Social Networks} ~\cite{Sukthankar-SocialInfo2014}
introduced a synthetic network generator designed for cloning social network
statistics of an existing dataset. The network starts with a small number of
nodes, and new nodes are added until the network reaches the required number. It
has two basic parameters: homophily and link density. A high \emph{homophily}
value reflects that links have higher chances to be established among the nodes with the
same label; these labels can be considered as being equivalent to community
membership.

Attribute Synthetic Generator (ASG) is a network generator for reproducing the
node feature distribution of standard networks and rewiring the network to
preferentially connect nodes that exhibit a high feature similarity. The network
is initialized with a group of three nodes, and new nodes and links are added to
the network based on link density, homophily, and feature similarity. As new
nodes are created, their labels are assigned based on the prior label
distribution. After the network has reached the same number of nodes as the
original social media dataset, each node initially receives a random attribute
assignment. Then a stochastic optimization process is used to move the initial
assignments closer to the target distribution extracted from social media
dataset using the Particle Swarm Optimization algorithm. The tuned attributes
are then used to add additional links to the network based on the feature
similarity parameter -- a source node is selected randomly and connected to the
most similar node. Multi-Link Generator (MLG) further  uses link co-occurrence statistics from the
original dataset to create a multiplex network. MLG uses the same network growth
process as ASG. Based on the link density parameter, either a new node is
generated with a label based on the label distribution of the target dataset or
a new link is created between two existing nodes.

\paragraph{LDBC SNB} Despite having a common Facebook-like dataset, thanks to three distinct workloads the Social Network Benchmark (SNB)~\cite{Erling:2015:LSN:2723372.2742786} provided by LDBC represents three distinct benchmarks. The network nodes correspond to people and the edges represent their friendship and messages they post in discussion trees on their forums. The three query workloads involve: (1) SNB-Interactive, i.e., complex read-only queries accessing a significant portion of data, (2) SNB-BI, i.e., queries accessing a high percentage of  entities and grouping them in various dimensions, and (3) SNB-Algorithms, involving graph analysis algorithms, such as community detection, PageRank, BFS, and clustering. The graph generator realizes power laws, uses skewed value distributions, and ensures reasonable correlations between graph structures and property values. To provide scalability it is implemented on top of Hadoop.

@@

>>>>

Another important aspect of social networks is the community effect. A detailed
study of structure of communities in 70 real-world networks is provided, e.g.,
in ~\cite{Leskovec:2008:SPC:1367497.1367591}.
~\cite{Prat-Perez:2014:CSS:2621934.2621942} analyzed the structure of
communities (clustering coefficient, triangle participation ratio, bridges,
diameter, conductance and size) in both real-world graphs and outputs of
existing graph generators such as LFR~\cite{PhysRevE.78.046110} and the
LDBC-SNB~\cite{Erling:2015:LSN:2723372.2742786}. They found out that discovered
communities  in different graphs have common distributions and that communities
in a single graph have diverse nature and are difficult to fit with a single
model.

@@

\paragraph{Realistic Social Network}
~\cite{Barrett:2009:GAL:1995456.1995598} focused on the construction of
realistic social networks. For this purpose the authors combine both public and
private data sets with large-scale agent-based techniques. The process works as
follows: In the first step it generates a synthetic data by combining public and
commercial databases. In the second step, it determines a set of activity
templates. A 24-hour activity sequence including geolocations is assigned to
each synthetic individual. To demonstrate the approach, the authors create a
synthetic US population consisting of people and households together with
respective geolocations. For this purpose the authors combine simulation and
data fusion techniques utilizing various real-world data sources such as U.S.
Census data, responses to an activity survey or a time-use survey. 
%Demographic information for each person and location, a minute-by-minute
%schedule of each person's activities, and the locations where these activities
%take place is generated by a combination of simulation and data fusion
%techniques.
The result is captured by a dynamic social contact network. Similar methods for
agent-based strategies have been reported
in~\cite{Bernstein:2013:SAS:2499604.2499609}.

\paragraph{Linkage vs. Activity Graphs} 
\cite{Yao2011} distinguished between two types of social network graphs -- the
\emph{linkage graph}, where nodes represent people and edges correspond to their
friendships, and the \emph{activity graph}, where nodes also represent people
but edges correspond to their interactions. On the basis of the analysis of
Flickr\footnote{\url{https://www.flickr.com/}} social links and
Epinions\footnote{\url{http://www.epinions.com/}} network of user interactions,
the authors discover that they both exhibit high clustering coefficient
(community structure), small diameter, and power-law degree distribution.
Considering the dynamic properties they both have relatively stable clustering
coefficient over time and follow the densification law. On the other hand,
diameter shrinking is not observed in Epinions activity graph and there is a
difference in degree correlation (i.e., frequency of mutual connections of
similar nodes) -- linkage graphs have positive degree correlation, activity
graphs have neutral degree correlation. With regards to the findings, the
proposed generator focusses on linkage graphs with positive degree correlation.
For this purpose it extends the forest fire spreading process
algorithm~\cite{Leskovec:2005:GOT:1081870.1081893} with link symmetry. It has
two parameters: the \emph{burning probability} $P_b$ and the \emph{symmetry
probability} $P_s$. $P_b$ ensures a BFS-based forward burning process in which
fire burns strongly with $P_b$ approaching 1.  $P_s$ ensures backward linking
from old nodes to new nodes and ``adds fuel to the fire as it brings more
links''. %It gives chances for big nodes to connect back to big nodes.

@@

\paragraph{Cloning of Social Networks} ~\cite{Sukthankar-SocialInfo2014}
introduces two synthetic generators to reproduce two characteristics typically
observed in social networks: node features and multiple link types. Both
generators etend on the generator proposed by~\cite{wang2011leveraging}.
which starts with a small number of
nodes, and new nodes are added until the network reaches the required number. It
has two basic parameters: homophily and link density. A high \emph{homophily}
value reflects that links have higher chances to be established among the nodes with the
same label; these labels can be considered as being equivalent to community
membership.

The first proposed generator is Attribute Synthetic Generator (ASG), used for
reproducing the node feature distribution of standard networks and rewiring the
network to preferentially connect nodes that exhibit a high feature similarity.
The network is initialized with a group of three nodes, and new nodes and links
are added to the network based on link density, homophily, and feature
similarity. As new nodes are created, their labels are assigned based on the
prior label distribution. After the network has reached the same number of nodes
as the original social media dataset, each node initially receives a random
attribute assignment. Then a stochastic optimization process is used to move the
initial assignments closer to the target distribution extracted from social
media dataset using the Particle Swarm Optimization algorithm. The tuned
attributes are then used to add additional links to the network based on the
feature similarity parameter -- a source node is selected randomly and connected
to the most similar node. The second proposed generator is Multi-Link Generator
(MLG) further  uses link co-occurrence statistics from the original dataset to
create a multiplex network. MLG uses the same network growth process as ASG.
Based on the link density parameter, either a new node is generated with a label
based on the label distribution of the target dataset or a new link is created
between two existing nodes.

\paragraph{LDBC SNB} Despite having a common Facebook-like dataset, thanks to
three distinct workloads the Social Network Benchmark
(SNB)~\cite{Erling:2015:LSN:2723372.2742786} provided by LDBC represents three
distinct benchmarks. The network nodes correspond to people and the edges
represent their friendship and messages they post in discussion trees on their
forums. The three query workloads involve: (1) SNB-Interactive, i.e., complex
read-only queries accessing a significant portion of data, (2) SNB-BI, i.e.,
queries accessing a high percentage of  entities and grouping them in various
dimensions, and (3) SNB-Algorithms, involving graph analysis algorithms, such as
community detection, PageRank, BFS, and clustering. The graph generator, called
Datagen, is a fork of S3G2 ~\cite{Pham2013} and realizes power laws, uses skewed
value distributions, and ensures reasonable correlations between graph
structures and property values. Additionally, it extends S3G2 with "spiky"
patterns in the distribution of social network activity along the timeline, also
provides the ability of generating update streams to the social network. Datagen
is also based on Hadoop in order to provide scalability, but comparet to S3G2 it
contains numerous performance improvements and the ability to be deterministic
regardless of the number of computer nodes used for the generation of the graphs
and for a given set of configuration parameters.

@@

\paragraph{Strengths and Weaknesses of Social Network Generators}
Compared to more general graph generators, social network generators focus
mainly on reproducing intra- and inter-node feature correlations. 
Among existing genrators, LDBC-SNB and S3G2 look
like the most advanced ones in terms of the complexity of the generated graph
and the amount of features and correlations they can generate, while providing a
large degree of scalability. Their generation process is based on input dictionaries and have configuration files that
allow tweaking parameters many parameters of the generated graphs, but their schema is mainly
static and cannot be easilly configured to meet the needs of other uses cases
besides the benchmarks they have been designed for. In this regard, the approaches
like those proposed in ~\cite{Nettleton2016} and ~\cite{Sukthankar-SocialInfo2014} offer a
more flexible and understandable configuration process to tweak the types,
values and correlations between different features.

Regarding the correlation between the underlying graph structure and the node
features, approaches such as LDBC-SNB, S3G2 or ~\cite{Nettleton2016} 
take into account this aspect and the generated graphs have ralistic
structural properties while similar nodes have a larger probability of being
connected. However, their approach seems to be more based on
intuition and common sense than to be backed up by an any study of how the
relation between structure and attributes showcase in real social networks. In
this reagard, this remains as a clear open challenge for social network
generators.

Finally, scalability is another aspect to be considered in social network graph
generators. LDBC-SNB and S3G2 are engineered with this in mind, thus they
provide a way to scale to billions of nodes and edges. This is not the case for
the other generators, which can make them impractical if our goal is to generate
real sized social network graphs.
